---
title: "HW6"
author: "Holly Finertie - HF2379"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(modelr)
```


## Problem 1 

### a) Load and tidy dataset: 

```{r}
#load and tidy dataset
birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = factor(recode(babysex, 
              `1` = "Male", 
              `2` = "Female")), 
    frace = factor(recode(frace, 
              `1` = "White", 
              `2` = "Black",
              `3` = "Asian", 
              `4` = "Puerto Rican", 
              `8` = "Other", 
              `9` = "Unknown")), 
    malform = factor(recode(malform, 
              `0` = "Absent", 
              `1` = "Present")), 
    mrace = factor(recode(mrace, 
              `1` = "White", 
              `2` = "Black",
              `3` = "Asian", 
              `4` = "Puerto Rican", 
              `8` = "Other"))) 

#check for missing values
anyNA(birthweight)
```

### b) Build a model:  
Many factors, including biological and social, affect birthweight. Per [March of Dimes](https://www.marchofdimes.org/complications/low-birthweight.aspxl), social factors that affect low birthweight include substances use, SES, violence, and race. Using the variables available in the `birthweight.csv` dataset, I hypothesize that family income, mother's race, and number of cigarettes smoked per day during pregnancy will be significantly associated with birthweight. 

First, I will test the bivariate associations between birthweight and each covariate of interest. Covariates significant at p-value <0.25 will be considered in the multinomial multivariable analysis. 

#### Bivariate Analysis

```{r}
bwt_model = birthweight %>% 
  mutate(
    mrace = fct_infreq(mrace)) %>% 
  select(bwt, fincome, gaweeks, mrace, smoken)

model_fincome = lm(bwt ~ fincome, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) 
```

* **Birthweight and Family Income:** At the 5% level of significant, family income is significantly associated with birthweight. For every $100 increase in monthly income, birthweight increases by `r model_fincome %>% filter(term == "fincome") %>% pull(estimate) %>% round(., digits = 2)` grams. 


```{r}
model_mrace = lm(bwt ~  mrace, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) 
```

* **Birthweight and Mother's Race:** At the 5% level of significant, mother's race is significantly associated with birthweight. Compared to white mothers, all other races have baby's with a lower birthweight. Black mothers have baby's who weigh `r model_mrace %>% filter(term == "mraceBlack") %>% pull(estimate) %>% round(., digits = 2)` grams less. Puerto Rican mothers have baby's who weigh `r model_mrace %>% filter(term == "mracePuerto Rican") %>% pull(estimate) %>% round(., digits = 2)` grams less. Asian mothers have baby's who weigh `r model_mrace %>% filter(term == "mraceAsian") %>% pull(estimate) %>% round(., digits = 2)` grams less. 


```{r}
model_smoken = lm(bwt ~  smoken, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* **Birthweight and Smoking:** At the 5% level of significant, smoking is significantly associated with birthweight. For every 1 additional cigarette smoked per day, birthweight decreases `r model_smoken %>% filter(term == "smoken") %>% pull(estimate) %>% round(., digits = 2)` grams. 


#### Multivariable Analysis

Now that we confirmed all three variables are separately and significantly associated with the outcome (birthweight), we can move forward with backwards step-wise analysis. I will fit a model with all three variables. If one is not signifcant at a 5% alpha level, I will remove that variable and repeat until all variables are significant. 


```{r}
model_social = lm(bwt ~ fincome + mrace + smoken, data = bwt_model) 

tidy_social = model_social %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

tidy_social
```

All three variables are significant at a 5% alpha level. 

```{r}
added_social = bwt_model %>% 
  modelr::add_residuals(model_social) %>% 
  modelr::add_predictions(model_social)

plot_social = added_social %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()

```

### c) Test Against Other Models: 

Based on the below plot, the model with the interaction terms (`bwt = bhead + blength + babysex + bhead x blength + bhead x babysex + blength x babysex + bhead x blength x babysex`) is the best fitting model with the lowest mean squared errors. 

```{r}
model_main_effects = lm(bwt ~ blength + gaweeks, data = birthweight) 

model_interaction = lm(bwt ~ bhead + blength + babysex 
                       + bhead*blength + bhead*babysex + blength*babysex
                       + bhead*blength*babysex, data = birthweight) 

bwt_cv = 
  crossv_mc(birthweight, 100) 

bwt_cv %>% pull(train) %>% .[[1]] %>% as_tibble

bwt_cv %>% pull(test) %>% .[[1]] %>% as_tibble

bwt_cv =
  bwt_cv %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))


bwt_cv = 
  bwt_cv %>% 
  mutate(
    model_social = map(train, ~ lm(bwt ~ fincome + mrace + smoken, data = bwt_model)),  
    model_main_effects = map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight)), 
    model_interaction = map(train, ~lm(bwt ~ bhead + blength + babysex 
                       + bhead*blength + bhead*babysex + blength*babysex
                       + bhead*blength*babysex, data = birthweight))) %>% 
  mutate(rmse_social = map2_dbl(model_social, test, ~rmse(model = .x, data = .y)),
         rmse_main_effects = map2_dbl(model_main_effects, test, ~rmse(model = .x, data = .y)),
         rmse_interaction = map2_dbl(model_interaction, test, ~rmse(model = .x, data = .y)))


bwt_cv %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = "Prediction Error Distributions Between Models"
  )
```

## Problem 2  

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())


weather_bootstrap = weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy))%>% 
  select(results) %>% 
  unnest(results) %>% 
  filter(term == "tmin")

weather_bootstrap %>% 
  ggplot(aes(x = estimate)) + 
  geom_density()

```


Across the the 5,000 estimates, the mean estimate is `r weather_bootstrap %>% summarize(mean = round(mean(estimate), digits = 2)) %>% pull(mean)` with standard error `r weather_bootstrap %>% summarize(mean = round(mean(std.error), digits = 2)) %>% pull(mean)`. Just looking at it visually, there doesn't appear to be any skewness and the distribution centers around 1.04. 












