---
title: "HW6"
author: "Holly Finertie - HF2379"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(modelr)
```


## Problem 1 

#### Load and tidy dataset: 

```{r}
#load and tidy dataset
birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = factor(recode(babysex, 
              `1` = "Male", 
              `2` = "Female")), 
    frace = factor(recode(frace, 
              `1` = "White", 
              `2` = "Black",
              `3` = "Asian", 
              `4` = "Puerto Rican", 
              `8` = "Other", 
              `9` = "Unknown")), 
    malform = factor(recode(malform, 
              `0` = "Absent", 
              `1` = "Present")), 
    mrace = factor(recode(mrace, 
              `1` = "White", 
              `2` = "Black",
              `3` = "Asian", 
              `4` = "Puerto Rican", 
              `8` = "Other"))) 

#check for missing values
anyNA(birthweight)
```

#### Build a model:  
Many factors, including biological and social, affect birthweight. Per [March of Dimes](https://www.marchofdimes.org/complications/low-birthweight.aspxl), social factors that affect low birthweight include substances use, SES, violence, and race. Using the variables available in the `birthweight.csv` dataset, I hypothesize that family income, mother's race, and number of cigarettes smoked per day during pregnancy will be significantly associated with birthweight. 

First, I will test the bivariate associations between birthweight and each covariate of interest. Covariates significant at p-value <0.25 will be considered in the multinomial multivariable analysis. 

##### Bivariate Analysis

```{r}
bwt_model = birthweight %>% 
  mutate(
    mrace = fct_infreq(mrace)) %>% 
  select(bwt, fincome, gaweeks, mrace, smoken)

model_fincome = lm(bwt ~ fincome, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) 
```

* **Birthweight and Family Income:** At the 5% level of significant, family income is significantly associated with birthweight. For every $100 increase in monthly income, birthweight increases by `r model_fincome %>% filter(term == "fincome") %>% pull(estimate) %>% round(., digits = 2)` grams. 


```{r}
model_mrace = lm(bwt ~  mrace, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) 
```

* **Birthweight and Mother's Race:** At the 5% level of significant, mother's race is significantly associated with birthweight. Compared to white mothers, all other races have baby's with a lower birthweight. Black mothers have baby's who weigh `r model_mrace %>% filter(term == "mraceBlack") %>% pull(estimate) %>% round(., digits = 2)` grams less. Puerto Rican mothers have baby's who weigh `r model_mrace %>% filter(term == "mracePuerto Rican") %>% pull(estimate) %>% round(., digits = 2)` grams less. Asian mothers have baby's who weigh `r model_mrace %>% filter(term == "mraceAsian") %>% pull(estimate) %>% round(., digits = 2)` grams less. 


```{r}
model_smoken = lm(bwt ~  smoken, data = bwt_model) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* **Birthweight and Smoking:** At the 5% level of significant, smoking is significantly associated with birthweight. For every 1 additional cigarette smoked per day, birthweight decreases `r model_smoken %>% filter(term == "smoken") %>% pull(estimate) %>% round(., digits = 2)` grams. 


##### Multivariable Analysis

Now that we confirmed all three variables are separately and significantly associated with the outcome (birthweight), we can move forward with backwards step-wise analysis. I will fit a model with all three variables. If one is not signifcant at a 5% alpha level, I will remove that variable and repeat until all variables are significant. 


```{r}
model_social = lm(bwt ~ fincome + mrace + smoken, data = bwt_model) 

tidy_social = model_social %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

tidy_social
```

All three variables are significant at a 5% alpha level. 


```{r}
added_social = bwt_model %>% 
  modelr::add_residuals(model_social) %>% 
  modelr::add_predictions(model_social)

plot_social = added_social %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()

```
